---
title: "Statistical Inference Assignement"
author: "François Grolleau"
date: "26/06/2020"
output: html_document
---

### Loading relevant packages

```{r setup, results=FALSE}
library(ggplot2)
library(datasets)
```

# PART 1

### Create matrix of dimention 1000 X 40
It's populated with random variable from exponetial distribution of lambda 0.2 
```{r, results=FALSE}
lambda<-0.2
samplesize<-40
n_sim<-1000
df<-matrix(rexp(samplesize*n_sim, lambda), n_sim, samplesize)
```

### Sampling distribution of 1000 sample means
Each sample was of size 40
vertical line stands for the theoretical mean mu
as a result of the central limit theorem the sampling distribution of the sample mean is normally distributed with mean = mu and sandart deviation = sigma/sqrt(n)
```{r, results=TRUE}
means<-apply(df, 1, mean)
hist(means)
abline(v=1/lambda, lwd=10)
```

### Sampling distribution of 1000 sample variances
Each sample was of size 40
vertical line stands for the theoretical variance sigma^2
as a result of the central limit theorem the sampling distribution of the sample variance is normally distributed with mean = sigma^2. 
The normality of the distribution is only approximated because of
i) the great kurtosis of the distrubution we are sampling from
and
ii) the limited sample size of each sample (40)
```{r, results=TRUE}
variances<-(apply(df, 1, sd))^2
hist(variances)
abline(v=1/lambda^2, lwd=10)
```

### Reshaping the data
So that ggplot can exploit it
```{r, results=TRUE}
df2<-data.frame(name=factor(rep(c("1000 exp", "1000 means of 40 exps", "1000 variances of 40 exps"), each=1000)) , value=c(df[,1], means, variances))
head(df2)
tail(df2)
```

### Calculating the relevant statistics
```{r, results=TRUE}
summarystat<-data.frame(distribution_mean=1/lambda,
                        mean_of_1000_means=mean(df2$value[which(df2$name=="1000 means of 40 exps")]),
                        mean_of_1000_variances=mean(df2$value[which(df2$name=="1000 variances of 40 exps")]))
summarystat 
```

### GGploting around
Plotting the distribution with corresponding names
```{r, results=TRUE}
g<-ggplot(df2, aes(value, color=name, fill=name)) + geom_histogram(linetype="blank", position="identity", alpha=.6, binwidth = 1)
g
```

Adding vertical lines of the means of the sampling distribution of 
i) the sample mean
ii) the sample variance

They are unbiased estimates of the sample mean and the sample variance respectively
```{r, results=TRUE}
g<-g + geom_vline(xintercept = summarystat$mean_of_1000_means)
g<-g + geom_vline(xintercept = summarystat$mean_of_1000_variances)
g
```

# PART 2
### Basic summary of the data
```{r, results=TRUE}
head(ToothGrowth)
summary(ToothGrowth)
str(ToothGrowth)
```

Putting dose as a factor and then describing the number of occurences of each factor (dose first, then supp)
```{r, results=TRUE}
ToothGrowth$dose<-factor(ToothGrowth$dose)
table(ToothGrowth$dose)
table(ToothGrowth$supp)
```

### GGploting around
# Distribution of len relative to supp
```{r, results=TRUE}
ggplot(ToothGrowth, aes(len, col=supp)) + geom_boxplot() + coord_flip()
```

t test assuming uniequal variances
95% CI an P-value are provided
```{r, results=TRUE}
t.test(ToothGrowth$len[which(ToothGrowth$supp=="OJ")],
       ToothGrowth$len[which(ToothGrowth$supp=="VC")], var.equal = FALSE)
```


# Distribution of len relative to doses
```{r, results=TRUE}
ggplot(ToothGrowth, aes(len, col=dose)) + geom_boxplot() + coord_flip()
```

t tests assuming uniequal variances
95% CIs an P-values are provided

```{r, results=TRUE}
t.test(ToothGrowth$len[which(ToothGrowth$dose==.5)],
       ToothGrowth$len[which(ToothGrowth$dose==1)])

t.test(ToothGrowth$len[which(ToothGrowth$dose==1)],
       ToothGrowth$len[which(ToothGrowth$dose==2)])

t.test(ToothGrowth$len[which(ToothGrowth$dose==.5)],
       ToothGrowth$len[which(ToothGrowth$dose==2)])
```
